---
title: "Coursera Practical Machine Learning Course Project- Prediction Assignment Writeup"
output: 
  html_document: 
    keep_md: yes
---

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load the required libraries:

```{r, results='hide', message=FALSE, warning=FALSE}
library(caret)
library(rpart)
library(rattle)
library(randomForest)
```

#### Download the data:

```{r}
rm(list = ls())
if (!file.exists("pml-training.csv")) {
  download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = "pml-training.csv")
}
if (!file.exists("pml-testing.csv")) {
  download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "pml-testing.csv")
}
training  <- read.csv("pml-training.csv", sep = ",", na.strings = c("", "NA"))
testing  <- read.csv("pml-testing.csv", sep = ",", na.strings = c("", "NA"))
```

### Split the training set:

We split the training data set into two data sets, 60% for training and 40% for testing.

```{r}
set.seed(98765)
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
myTraining <- training[inTrain, ]
myTesting <- training[-inTrain, ]
dim(myTraining); dim(myTesting)
```

### Data cleaning:

Lets see what NZV features we have.

```{r}
myDataNZV <- nearZeroVar(myTraining, saveMetrics=TRUE)
```

```{r}
myNZVvars <- names(myTraining) %in% rownames(myDataNZV[myDataNZV$nzv == "TRUE", ])
myTraining <- myTraining[!myNZVvars]
myTraining <- myTraining[c(-1)]
dim(myTraining)
```

We then remove the features with more than 60% NA values.

```{r}
trainingV3 <- myTraining
for(i in 1:length(myTraining)) { 
        if( sum( is.na(myTraining[, i] ) ) / nrow(myTraining) >= .6 ) { 
        for(j in 1:length(trainingV3)) {
            if( length(grep(names(myTraining[i]), names(trainingV3)[j]) ) == 1) {
                trainingV3 <- trainingV3[ , -j]
            }   
        } 
    }
}

myTraining <- trainingV3
dim(myTraining)
```

```{r, include=FALSE}
clean1 <- colnames(myTraining)
clean2 <- colnames(myTraining[, -58]) 
myTesting <- myTesting[clean1]
testing <- testing[clean2]

dim(myTesting)
```

```{r, include=FALSE}
for (i in 1:length(testing) ) {
        for(j in 1:length(myTraining)) {
        if( length( grep(names(myTraining[i]), names(testing)[j]) ) == 1)  {
            class(testing[j]) <- class(myTraining[i])
        }      
    }      
}
#To make sure Coertion works, we use the following technique:
testing <- rbind(myTraining[2, -58] , testing)
testing <- testing[-1,]
```

### Machine learning and Decision Tree:

Predict the diferent types of exercises with a simple decision tree.

```{r}
modFitA1 <- rpart(classe ~ ., data=myTraining, method="class")
```

This is what the decision tree looks like:

```{r}
fancyRpartPlot(modFitA1)
```

And here are the results:

```{r}
predictionsA1 <- predict(modFitA1, myTesting, type = "class")
confusionMatrix(predictionsA1, myTesting$classe)
```

Now we predict with 500 trees using the default randomForest function:

```{r}
modFitB1 <- randomForest(classe ~. , data=myTraining)
predictionsB1 <- predict(modFitB1, myTesting, type = "class")
```

Take a look at the results obtained with randomForest:

```{r}
confusionMatrix(predictionsB1, myTesting$classe)
```
